### 2017-05-23 sections/implementation.tex
Update on Overleaf.

\noindent
where $e$ is NRMSD used as an error limit that can not be exceeded to achieve reasonable quality of approximation.

@@

\begin{itemize}
\item The algorithm is most efficient in cases where one dimension of the input matrix is very low compared to the other. However, this is almost always the case when compressing results from FEM -- number of time steps seldom exceeds hundreds.
\item Moreover, time steps can be split into smaller ranges and the algorithm can be applied on each range separately. This will improve performance and can also increase quality of compression if the key time steps on the range boundaries are carefully selected.
\item Randomized SVD algorithm has the same order of algorithmic complexity but yet can significantly reduce execution time. (The implementation of the particular randomized SVD algorithm that was used is described in \cite{Halko2011}.)
\end{itemize}

>>>>

\noindent
where $e$ is $NRMSD$ used as an error limit that can not be exceeded to achieve reasonable quality of approximation.

@@

\begin{itemize}
\item The algorithm is most efficient in cases where one dimension of the input matrix is very low compared to the other. However, this is almost always the case when compressing results from FEM -- number of time steps seldom exceeds hundreds.
\item Moreover, time steps can be split into smaller ranges and the algorithm can be applied on each range separately. This will improve performance and can also increase quality of compression if the key time steps on the range boundaries are carefully selected.
\item Randomized SVD algorithm has the same order of algorithmic complexity when full decomposition is required, but yet can significantly reduce execution time. However, the benchmarks are not designed to highlight the benefits of randomized SVD algorithms. The main advantage of the randomized SVD is in the ability to choose the rank of the approximation matrix in advance. In that case only limited number of singular values and corresponding singular vectors is calculated and algorithm performs much faster.
\end{itemize}

### 2017-05-23 sections/math.tex
Update on Overleaf.

% Singular Value Decomposition in general
% TODO: Describe how to calculate decomposition of arbitrary matrix with example

@@

Exact SVD of a $m \times n$ matrix has computational complexity $O(min(mn^2, m^2n))$ using the "big-O" notation. In \cite{Holmes2007, Candes2011, Woolfe2008, Martinsson2011} are described randomized algorithms for constructing approximate matrix factorizations that have better asymptotic complexity. Their main purpose is to accelerate calculations of matrix decompositions which tend to be very slow for large data sets.

@@

%\begin{equation}
%\mtrx{A}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{n}}S_{n}\mtrx{V_{n}^{T}},
%\label{eq:svd-expansion}
%\end{equation}

@@

%\begin{equation}
%\mtrx{A'}=\mtrx{U_{1}}S_{1}\mtrx{V_{1}^{T}}+\mtrx{U_{2}}S_{2}\mtrx{V_{2}^{T}}+ ... +\mtrx{U_{r}}S_{r}\mtrx{V_{r}^{T}}.
%\label{eq:svd-approx-expansion}
%\end{equation}

@@

% subsection: Sparse matrix of details: Mozna uplne vyhodit, stejne to neni implementovano -- uvest jen v sekci Optimization nebo v Conclusion

>>>>

% Singular Value Decomposition in general

@@

@@

@@

@@

\subsection{Randomized SVD}

Exact SVD of a $m \times n$ matrix has computational complexity $O(min(mn^2, m^2n))$ using the "big-O" notation. When applied on large data sets it tends to be very time-consuming. In \cite{Holmes2007, Candes2011, Woolfe2008, Martinsson2011} are described randomized methods for constructing approximate matrix factorizations which offer significant speedups over classical methods.

As the implementation of the randomized SVD algorithm \textit{RedSVD} library was used. \textit{RedSVD} is an open-source software package for solving several matrix decompositions. The particular implementation of randomized version of decomposition is based on the algorithm described in \cite{Halko2011}. The brief summary of the algorithm steps follows below.

% redsvd algorithm description
Let $\mtrx{A}$ be a matrix to be analyzed with $n$ rows and $m$ columns, and $r$ be the rank of a truncated SVD (if $r = min(n, m)$ is chosen, then it is the original SVD).

At first, a random Gaussian matrix $\mtrx{O}$ with $m$ rows and $r$ columns is created and a sample matrix $\mtrx{Y} = \mtrx{A}^{\mathsf{T}} \mtrx{O}$ is computed. Then the Gram-Schmidt process is applied to $\mtrx{Y}$ so that each column of $\mtrx{Y}$ is orthonormalized. Then matrix $\mtrx{B} = \mtrx{A}\mtrx{Y}$ with $n$ rows and $r$ columns is computed. Although the size of $\mtrx{Y}$ is much smaller than that of $\mtrx{A}$, $\mtrx{Y}$ holds the information of $\mtrx{A}$; that means $\mtrx{A}\mtrx{Y}\mtrx{Y}^{\mathsf{T}} \approx \mtrx{A}$. Intuitively, the row information is compressed by $\mtrx{Y}$ and can be decompressed by $\mtrx{Y}^{\mathsf{T}}$.

Similarly, another random Gaussian matrix $\mtrx{P}$ with $r$ rows and $r$ columns is created, and matrix $\mtrx{Z} = \mtrx{B}\mtrx{P}$ is computed. As in the previous case, the columns of $\mtrx{Z}$ are orthonormalized by the Gram-Schmidt process and $\mtrx{Z}\mtrx{Z}^{\mathsf{T}} \mtrx{B} \approx \mtrx{B}$. Then matrix $\mtrx{C} = \mtrx{Z}^{\mathsf{T}} \mtrx{B}$ is computed.

Finally, algorithm performs SVD of $\mtrx{C}$ using the traditional SVD solver, and obtains $\mtrx{C} = \mtrx{\widetilde{U}}\mtrx{\widetilde{S}}\mtrx{\widetilde{V}}^{\mathsf{T}}$ where $\mtrx{\widetilde{U}}$ and $\mtrx{\widetilde{V}}$ are orthonormal matrices, and $\mtrx{\widetilde{S}}$ is a diagonal matrix whose entries are singular values of matrix $\mtrx{C}$. Since matrix $\mtrx{C}$ is usually very small, computational time of this step is negligible.

Now $\mtrx{A}$ is decomposed as

\begin{eqnarray}
\mtrx{A} &\approx& \mtrx{A}\mtrx{Y}\mtrx{Y}^{\mathsf{T}} =
\\
&=& \mtrx{B}\mtrx{Y}^{\mathsf{T}} \approx \nonumber
\\
&\approx& \mtrx{Z}\mtrx{Z}^{\mathsf{T}}\mtrx{B}\mtrx{Y}^{\mathsf{T}} = \nonumber
\\
&=& \mtrx{Z}\mtrx{C}\mtrx{Y}^{\mathsf{T}} = \nonumber
\\
&=& \mtrx{Z}\mtrx{\widetilde{U}} \mtrx{\widetilde{S}} \mtrx{\widetilde{V}}^{\mathsf{T}}\mtrx{Y}^{\mathsf{T}} = \nonumber
\\
&=& \mtrx{Z}\mtrx{\widetilde{U}} \mtrx{\widetilde{S}} (\mtrx{Y}\mtrx{\widetilde{V}})^{\mathsf{T}}. \nonumber
\label{eq:redsvd}
\end{eqnarray}

\noindent
Both $\mtrx{Z}\mtrx{\widetilde{U}}$ and $\mtrx{Y}\mtrx{\widetilde{V}}$ are orthornormal, $\mtrx{Z}\mtrx{\widetilde{U}}$ contains approximations of left singular vectors, and $\mtrx{Y}\mtrx{\widetilde{V}}$ contains approximations of right singular vectors of matrix $\mtrx{A}$. $\mtrx{\widetilde{S}}$ contains approximations of the first $r$ singular values of matrix $\mtrx{A}$.

% http://mathoverflow.net/questions/161252/what-is-the-time-complexity-of-truncated-svd
%Exact SVD of a $m \times n$ matrix has time complexity $O(min(mn^2, m^2n))$ using the "big-O" notation. Assume the data points are in the columns of $\mtrx{A} \in \mtrx{M_{m,n}}(\mathbb{R})$ where $m \leq n$. Note that $\mtrx{AA^T}$ is the dataset covariance matrix. Then a simple method is to randomly choose $k<m$ columns of $\mtrx{A}$ that form a matrix $\mtrx{B}$. Statistically, the SVD of $\mtrx{BB^T}$ will be close to that of $\mtrx{AA^T}$; thus it suffices to calculate the SVD of $\mtrx{B}$, the complexity of which, is only $O(k^2m)$.

% subsection: Sparse matrix of details ?

### 2017-05-23 sections/results.tex
Update on Overleaf.

% MELA BY ZDE BYT DISKUZE, PROC PRI ROSTOUCIM POCTU RADKU SE NELINEARNE ZVYSUJE POCET OPRACI, ALE PRO ROSTOUCI POCET SLOUPCU SE POCET OPERACI ZVETSUJE LINEARNE.

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/executionTime_varyingRows}
\caption{Dependency of SVD execution time on number of rows of an input matrix.}
\label{fig:ExeTime_rows}
\end{figure}

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/executionTime_varyingColumns}
\caption{Dependency of SVD execution time on number of columns of an input matrix.}
\label{fig:ExeTime_columns}
\end{figure}

% reactor containment
Behavior of the compression strategy introduced is presented on three real world examples. First example is an analysis of aging of nuclear power plant's containment made from prestressed concrete. Finite element mesh used in this analysis is in Figure \ref{fig:temelin:mesh}. More details about the analysis can be found in references \cite{Kruis2012} and \cite{Koudelka2009}. This analysis includes high number of analysis time steps (thousands) with very little differences between them. There is therefore potential for compression ratio to be very high as proven in Figure \ref{fig:temelin:NRMSD} that examines the impact of changes in the compression ratio to the mean error of approximation.

>>>>

These results confirm that the SVD implementation has algorithm complexity $O(m^2n)$, where $m$ is number of rows, $n$ is number of columns, and $m < n$. In case of $m > n$ the complexity would be $O(mn^2)$ as the algorithm takes advantage of non-squareness in that its complexity is only linear in the greater dimension.

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/executionTime_varyingRows}
\caption{Dependency of SVD execution time on number of rows of an input matrix (having fixed number of columns).}
\label{fig:ExeTime_rows}
\end{figure}

\begin{figure}[H]
\centering\includegraphics[width=\textwidth]{figures/executionTime_varyingColumns}
\caption{Dependency of SVD execution time on number of columns of an input matrix (having fixed number of rows).}
\label{fig:ExeTime_columns}
\end{figure}

% reactor containment
Behavior of the compression strategy introduced is presented on three real world examples. First example is an analysis of aging of nuclear power plant's containment made from prestressed concrete. Finite element mesh used in this analysis is in Figure \ref{fig:temelin:mesh}. More details about the analysis can be found in \cite{Kruis2012} and \cite{Koudelka2009}. This analysis includes high number of analysis time steps (thousands) with very little differences between them. There is therefore potential for compression ratio to be very high as proven in Figure \ref{fig:temelin:NRMSD} that examines the impact of changes in the compression ratio to the mean error of approximation.
